---
title: "Project"
author: "Fatima Alsheikh"
date: "January 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

##Loading and preprocessing the data
The file wil be [downloaded](d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip), unzipped and saved in the directory. The data will be read as .csv file, the date varible class will be changed into Date and a column determing the weekday will be added.

```{r}
if(!file.exists("./Project")){dir.create("./Project")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./Project/repdata.zip")
unzip(zipfile="./Project/repdata.zip",exdir="./Project")
mydata<- read.csv("./Project/activity.csv")
head(mydata)
```

##What is mean total number of steps taken per day?
We will calculate the total number of steps taken per day, Make a histogram of the total number of steps taken each day and Calculate and report the mean and median of the total number of steps taken per day

```{r}
histdata<- aggregate(steps~ date, mydata, sum)
head(histdata)
hist(histdata$steps, breaks = 53, xlab = "Total number of steps taken each day", main = "Histogram of the total number of steps taken each day")
mean<- mean(histdata$steps)
median<- median(histdata$steps)
mean
median
```

##What is the average daily activity pattern?
We will Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) and determine the 5 min interval (on average) with maximum number of steps. 

```{r}
plotdata<- aggregate(steps~ interval, mydata, FUN = function(mydata) c(mean= mean(mydata))
)
plot(plotdata$interval, plotdata$steps, type = "l", xlab = "interval", ylab = "number of steps", main = "average daily activity pattern")
maximum<- plotdata[which.max(plotdata$steps),]
maximum
```

##Imputing missing values
The missing value will be imputed using mice package, and examine the difference before and after imputing
```{r results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(mice)
```
```{r}
newdata<- mice(mydata)
summary(newdata)
newdata<- complete(newdata)
histnewdata<- aggregate(steps~ date, newdata, sum)
par(mfrow = c(1,2))
hist(histnewdata$steps, breaks = 53, xlab = "Total number of steps taken each day", main = "Histogram of imputed data")
hist(histdata$steps, breaks = 53, xlab = "Total number of steps taken each day", main = "Histogram of original data")
mean1<- mean(histnewdata$steps)
median1<- median(histnewdata$steps)
mean1
median1

difference_in_mean<- mean- mean1
difference_in_median<- median- median1
difference_in_mean
difference_in_median
```

##Are there differences in activity patterns between weekdays and weekends?
```{r}
newdata$date<- as.Date(newdata$date)
newdata$weekday<- weekdays(newdata$date)
Weekends<- subset(newdata, weekday==c("Saturday", "Sunday"))
Weekdays<- subset(newdata, weekday!=c("Saturday", "Sunday"))
plotweekends<- aggregate(steps~ interval, Weekends, FUN = function(Weekends) c(mean= mean(Weekends)))
plotweekdays<- aggregate(steps~ interval, Weekdays,  FUN = function(Weekdays) c(mean= mean(Weekdays)))
par(mfrow= c(1,2))

plot(plotweekdays$interval, plotweekdays$steps, type = "l", xlab = "interval", ylab = "number of steps", main = "Weekdays")
plot(plotweekends$interval, plotweekends$steps, type = "l", xlab = "interval", ylab = "number of steps", main = "Weekends")    
```